Missing values

### 2022-02-10 Trilemma_clustering.Rmd
updating figures and missing values analysis

The bottle neck is certaintly our proxies of inequality, no matter what we use. Below a mock up example with some arbitrary choices.

@@

dat <- ecofp %>%
  filter(record == "EFConsPerCap") %>%
  select(countryName, isoa2, year, value) %>% 
  rename(EFconsPerCap = value, country = countryName) %>% 
  filter(year >= 1990)

dat <- dat %>% 
  left_join(
    .,
    inq %>% 
      filter(!is.na(gini_reported), year >= 1990) %>% 
      select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
      group_by(country, year, isoa2, country_code) %>% 
      summarize(gini_mean = mean(gini_reported, na.rm = TRUE))

@@

dat <- dat %>% 
  left_join(
    ., 
    dat1 %>% 
      select(-last_col()) %>% 
      filter(indicator_name == "GNI per capita (US$)") %>% 
      pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>%
      mutate(year = str_remove(year, "x")) %>% 
      mutate(year = as.numeric(year)) %>% 
      mutate(ismissing = is.na(value)) %>% 
      filter(year >= 1990) %>% 
      rename(country = country_name, gni = value)
  )

@@

key_countries <- missingness %>% 
    filter(missingness < 0.3) %>% 
    pull(country)
```

@@

```{r message=FALSE, warning=FALSE}
df_dat %>% 
    select(year, EFconsPerCap, gni2, gini) %>% 
    GGally::ggpairs(
        columns = 3:5 
    # lower = list(continuous = GGally::wrap("points", alpha = 0.5, size = 0.5)),
    # diag = list(continuous = GGally::wrap("densityDiag", alpha = 0.2)),
    # upper = list(continuous = GGally::wrap("cor", size = 2))
    ) + theme_light(base_size = 20)
```

@@

```{r}
p1 <- df_dat %>% 
  ggplot(aes(EFconsPerCap, gini, group = country)) +
  geom_point(size = 0.5) +
  geom_path(size = 0.2, aes(color = year)) +
  theme(legend.position = c(0.8, 0.8))

p2<- df_dat %>% 
  ggplot(aes(gni2, gini, group = country)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  geom_path(size = 0.2, aes(color = year)) +
  scale_x_log10()

@@

```{r}
df_dat |> 
    as_tibble() |> 
    ungroup() |> 
    group_by(country) |> 
    summarize(mgt = sum(magni, na.rm = TRUE), 
              theta = sum(theta, na.rm = TRUE),
              gamma = sum(gamma, na.rm = TRUE)) |> 
    arrange(desc(mgt))
```
```{r}
df_dat |> 
    as_tibble() |> 
    select(country, year, d_EF, d_gini, d_gni) |> 
    pivot_longer(cols = starts_with("d_"), names_to = "vars", values_to = "vals") |> 
    filter(!is.na(vals)) |> 
    ggplot(aes(year, vals, group = country)) +
    geom_path(size = 0.2, alpha = 0.5) +
    facet_wrap(~vars, scales = "free") +
    theme_light(base_size = 18)
```
```{r}
df_dat |> 
    ggplot(aes(EFconsPerCap, gini)) +
    geom_density2d() +
    geom_point(aes(color = gni_log)) +
    geom_path(aes(group = country), size = 0.15) +
    scico::scale_color_scico() +
    theme_light(base_size = 18)
```
```{r}
df_dat |> 
    ggplot(aes(gni_log, gini)) +
    geom_density2d() +
    geom_point(aes(color = EFconsPerCap)) +
    geom_path(aes(group = country), size = 0.15) +
    scico::scale_color_scico(palette = "berlin") +
    theme_light(base_size = 18)
```
```{r}
df_dat |> 
    ggplot(aes(gni_log, EFconsPerCap)) +
    geom_density2d() +
    geom_point(aes(color = gini)) +
    geom_path(aes(group = country), size = 0.15) +
    scico::scale_color_scico() +
    theme_light(base_size = 18)
```
Who is the weird trajectory out there, world? No it's Luxemburg

>>>>

The bottle neck is certaintly our proxies of inequality, no matter what we use. Based on the missing data explorations, inequality data improves from 1995 onward.

@@

dat <- ecofp %>%
  filter(record == "EFConsPerCap") %>%
  select(countryName, isoa2, year, value) %>% 
  rename(EFconsPerCap = value, country = countryName) %>% 
  filter(year >= 1995)

dat <- dat %>% 
  left_join(
    .,
    inq %>% 
      filter(!is.na(gini_reported), year >= 1995) %>% 
      select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
      group_by(country, year, isoa2, country_code) %>% 
      summarize(gini_mean = mean(gini_reported, na.rm = TRUE))

@@

dat <- dat %>% 
  left_join(
    ., 
    dat1 %>% 
      select(-last_col()) %>% 
      filter(indicator_name == "GNI per capita (US$)") %>% 
      pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>%
      mutate(year = str_remove(year, "x")) %>% 
      mutate(year = as.numeric(year)) %>% 
      mutate(ismissing = is.na(value)) %>% 
      filter(year >= 1995) %>% 
      rename(country = country_name, gni = value)
  )

@@

key_countries <- missingness %>% 
    filter(missingness <= 0.3) %>% 
    pull(country)

droped_countries <- missingness %>% 
    filter(missingness > 0.3) %>% 
    pull(country)
```

@@

```{r message=FALSE, warning=FALSE}
p <- df_dat %>% 
    select(year, `Ecological footprint` = EFconsPerCap, GNI = gni2, Gini = gini) %>% 
    GGally::ggpairs(
        columns = 3:5 ,
    lower = list(continuous = GGally::wrap("points", alpha = 0.5, size = 0.5)),
    diag = list(continuous = GGally::wrap("densityDiag", alpha = 0.2))
    # upper = list(continuous = GGally::wrap("cor", size = 2))
    ) + theme_light(base_size = 6)

# ggsave(
#     plot = p,
#     file = "correlogram.png", path = "figures/", device = "png",
#     width = 4, height = 4, dpi = 300, bg = "white"
# )
p + theme_light(base_size = 20)
```

@@

```{r}
p1 <- df_dat %>% 
    ggplot(aes(EFconsPerCap, gini, group = country)) +
    geom_point(size = 0.5) +
    geom_path(size = 0.2, aes(color = year)) +
    annotate(geom = "rect", xmin = 0, xmax = 1.68, ymin = 20, ymax = 30,
             fill = "skyblue", alpha = 0.25) +
    annotate(geom ="rect",xmin = 0, xmax = 1.68, ymin = 30, ymax = 62,
             fill = "orange", alpha = 0.1) +
    annotate(geom ="rect", xmin = 1.68, xmax = 18, ymin = 20, ymax = 30,
             fill = "orange", alpha = 0.1) +
    annotate(geom="rect", xmin = 1.68, xmax = 18, ymin = 30, ymax = 62,
             fill = "red", alpha = 0.1) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
    labs(y = "Gini", x = "Ecological footprint") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

p2<- df_dat %>% 
  ggplot(aes(gni2, gini, group = country)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  geom_path(size = 0.2, aes(color = year)) +
    annotate(geom = "rect", xmin = 10, xmax = 1.68, ymin = 20, ymax = 30,
             fill = "skyblue", alpha = 0.25) +
    annotate(geom ="rect",xmin = 0, xmax = 1.68, ymin = 30, ymax = 62,
             fill = "orange", alpha = 0.1) +
    annotate(geom ="rect", xmin = 1.68, xmax = 18, ymin = 20, ymax = 30,
             fill = "orange", alpha = 0.1) +
    annotate(geom="rect", xmin = 1.68, xmax = 18, ymin = 30, ymax = 62,
             fill = "red", alpha = 0.1) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 12746, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
      scale_x_log10() +
    labs(y = "Gini", x = "Ecological footprint") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

@@

```{r}
df_dat |> 
    as_tibble() |> 
    ungroup() |> 
    group_by(country) |> 
    summarize(mgt = sum(magni, na.rm = TRUE), 
              theta = sum(theta, na.rm = TRUE),
              gamma = sum(gamma, na.rm = TRUE)) |> 
    arrange(desc(mgt))
```
```{r fig.width=6, fig.height=3}
p <- df_dat |> 
    as_tibble() |> 
    select(country, year, d_EF, d_gini, d_gni) |> 
    pivot_longer(cols = starts_with("d_"), names_to = "vars", values_to = "vals") |> 
    filter(!is.na(vals)) |> 
    mutate(vars = case_when(
        vars == "d_EF" ~ "Ecological Footprint",
        vars == "d_gni" ~ "GNI [log]",
        vars == "d_gini" ~ "Gini"
    )) |> 
    ggplot(aes(year, vals, group = country)) +
    geom_path(size = 0.2, alpha = 0.5) +
    facet_wrap(~vars, scales = "free") +
    labs(x = "Year", y = "First difference") +
    theme_light(base_size = 7)

# ggsave(
#     plot = p, path = "figures/", file = "first_diff.png", device = "png",
#     width = 6, height = 3, dpi = 300, bg = "white")
p + theme_light(base_size = 20)
```
```{r fig.width=3, fig.height=3}
df_dat |> 
    ggplot(aes(EFconsPerCap, gini)) +
    geom_hline(yintercept = c(30,40), linetype = 2, color = "gray50") +
    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
    geom_density2d() +
    geom_point(aes(color = gni_log)) +
    geom_path(aes(group = country), size = 0.15) +
    scico::scale_color_scico("GNI [log]") +
    labs(y = "Gini", x = "Ecological footprint") +
    theme_light(base_size = 18) +
    theme(legend.position = "top", panel.grid = element_blank())
```
```{r}
df_dat |> 
    ggplot(aes(gni_log, gini)) +
    geom_density2d() +
    geom_point(aes(color = EFconsPerCap)) +
    geom_path(aes(group = country), size = 0.15) +
    scico::scale_color_scico(palette = "berlin") +
    theme_light(base_size = 18)
```
```{r}
df_dat |> 
    ggplot(aes(gni_log, EFconsPerCap)) +
    geom_density2d() +
    geom_point(aes(color = gini)) +
    geom_path(aes(group = country), size = 0.15) +
    scico::scale_color_scico() +
    theme_light(base_size = 18)
```
Who is the weird trajectory out there, world? No it's Luxemburg

### 2022-02-10 missing_values.Rmd
updating figures and missing values analysis

```{r}
ecofp %>%
    filter(record == "EFConsPerCap") %>%
    mutate(missing = is.na(value)) %>% 
    ggplot(aes(year, countryName)) +
    geom_tile(aes(fill = value)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```

```{r fig.height=3, fig.width=8}
g1 <- ecofp %>%
    filter(record == "EFConsPerCap") %>%
    select(countryName, year, value) %>% 
    pivot_wider(id_col = countryName, names_from = year, values_from = value) %>% 
    pivot_longer(cols = 2:last_col(), names_to = "year", values_to = "value") %>% 
    mutate(missing = is.na(value), year = as.numeric(year)) %>% 
    group_by(year) %>% 
    summarise(n_nas  = sum(missing)) %>% 
    ggplot(aes(year, n_nas)) + 
    geom_col(aes(fill = (year>1995)), alpha = 0.5) + 
    coord_flip() +
    theme_light(base_size = 7) +
    theme(legend.position = c(0.8,0.8)) + labs(title = "Missing values per year")

g2 <- ecofp %>%
    filter(record == "EFConsPerCap", year > 1995) %>%
    select(countryName, year, value) %>% 
    pivot_wider(id_col = countryName, names_from = year, values_from = value) %>% 
    pivot_longer(cols = 2:last_col(), names_to = "year", values_to = "value") %>% 
    mutate(missing = is.na(value), year = as.numeric(year)) %>% 
    group_by(countryName) %>% 
    summarise(n_nas  = sum(missing)) %>% 
    filter(n_nas > 0) %>% 
    mutate(countryName = as_factor(countryName) %>% fct_reorder(., n_nas)) %>% 
    ggplot(aes(countryName, n_nas)) + 
    geom_col(aes(fill = (n_nas < 11)), alpha = 0.5) + coord_flip() +
    theme_light(base_size = 7) + 
    labs(title = "Missing values per country after 1995") +
    theme(legend.position = c(0.8, 0.2))

g1 + g2
```

@@

```{r}
inq %>% 
    filter(year > 1950, resource_detailed == "Income, gross") %>% 
    ggplot(aes(gini_reported, palma)) +
    geom_miss_point(size = 1, alpha = 0.5) +
    theme(legend.position = c(0.1,0.8))
```

@@

```{r}
inq %>% 
    filter(year > 1950, resource_detailed == "Income, gross") %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini") %>% 
    mutate(year = as.numeric(year)) %>% 
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = gini)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```
```{r}
inq %>% 
    filter(!is.na(palma), year > 1990) %>% select(country, year, palma) %>% 
    group_by(country, year) %>% 
    summarize(palma_mean = mean(palma, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = palma_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "palma_mean") %>%
    mutate(year = as.numeric(year)) %>%
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = palma_mean)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```
To be fair with Gini, let's to the same over all Gini's reported. Even if Gini comes computed by different studies and in different currencies, the coefficient is always between 0-100, meaning it is comparable.

```{r}
inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year)) %>%
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = gini_mean)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```
So how many country / years can we use? Remember the data has `r inq$country %>% unique() %>% length()` countries

```{r}
inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year), missing = is.na(gini_mean)) %>% 
    group_by(year) %>% 
    summarise(n_nas = sum(missing), missingness = n_nas / 200)
```
We are dealing with ~50% of missing values on the best years, at best we can work with 100 countries. Which ones?

```{r}
inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year), missing = is.na(gini_mean)) %>% 
    group_by(country) %>% 
    summarise(n_nas = sum(missing), missingness = n_nas / (2018-1990)) %>% 
    arrange(missingness)
```

@@

@@

```{r}
dat1 %>% 
    select(-x33) %>% 
    filter(str_detect(indicator_name, "GNI")) %>% 
    pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>% 
    mutate(year = str_remove(year, "x")) %>% 
    mutate(year = as.numeric(year)) %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(color = country_name), show.legend = FALSE) +
    #geom_miss_point(show.legend = FALSE) +
    facet_wrap(~indicator_name, scales = "free_y")
```

```{r}
## How many missing values? 10506, but who has more?
missingness <- dat1 %>% 
    select(-x33) %>% 
    filter(str_detect(indicator_name, "GNI")) %>% 
    pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>% 
    mutate(year = str_remove(year, "x")) %>% 
    mutate(year = as.numeric(year)) %>% 
    mutate(ismissing = is.na(value)) %>% 
    group_by(country_code, indicator_name, country_name) %>% 
    summarize(missingness = sum(ismissing)/n())

missingness %>% 
    ggplot(aes(country_code, indicator_name)) +
    geom_tile(aes(fill = missingness)) + 
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.x = element_text(angle =90), legend.position = "top")
```

@@

```{r}
missingness %>% 
    ungroup() %>% 
    group_by(indicator_name) %>% 
    summarize(x = mean(missingness)) %>% 
    arrange(x)
```

```{r}
missingness %>% 
    ungroup() %>% 
    group_by(country_name) %>% 
    summarize(x = mean(missingness)) %>% 
    arrange(desc(x)) 
```
It seems that `GNI per capita (current US$)` is the one with less missing values. Tong recomments using the one with Atlas method wich is close on missingness, only 18%. A threshold of 50% missigness per country will left out places like Somalia, Iraq, Costa Rica, Zambia, China or Chile. Perhaps 60% is tolerable?

>>>>

```{r}
g0 <- ecofp %>%
    filter(record == "EFConsPerCap") %>%
    select(countryName, year, value) %>% 
    pivot_wider(id_col = countryName, names_from = year, values_from = value,
                values_fill = NA) %>% 
    pivot_longer(cols = 2:last_col(), names_to = "year", values_to = "value") %>% 
    mutate(missing = is.na(value), year = as.numeric(year)) %>% 
    ggplot(aes(year, countryName)) +
    geom_tile(aes(fill = value)) +
    scale_fill_viridis_c(
        "Ecological footprint",option = "D", na.value = "grey75")+
    labs(x = "Year", y = "Countries (N=211)", tag = "A") +
    scale_x_continuous(breaks = c( 1970, 1980, 1990, 2000, 2010)) +
    theme_light(base_size = 7)  +
    theme(axis.text.y = element_blank(), legend.position = "top")

g0
```

```{r fig.height=5, fig.width=7}
g1 <- ecofp %>%
    filter(record == "EFConsPerCap") %>%
    select(countryName, year, value) %>% 
    pivot_wider(id_col = countryName, names_from = year, values_from = value) %>% 
    pivot_longer(cols = 2:last_col(), names_to = "year", values_to = "value") %>% 
    mutate(missing = is.na(value), year = as.numeric(year)) %>% 
    group_by(year) %>% 
    summarise(n_nas  = sum(missing)) %>% 
    ggplot(aes(year, n_nas)) + 
    geom_col( alpha = 0.5) + 
    geom_vline(xintercept = 1995, color = "red", linetype = 2) +
    coord_flip() +
    theme_light(base_size = 7) +
    theme(legend.position = c(0.8,0.8)) + 
    labs(y = "Number of missing values", x="Year", tag = "B")
#g1
g2 <- ecofp %>%
    filter(record == "EFConsPerCap", year > 1995) %>%
    select(countryName, year, value) %>% 
    pivot_wider(id_col = countryName, names_from = year, values_from = value) %>% 
    pivot_longer(cols = 2:last_col(), names_to = "year", values_to = "value") %>% 
    mutate(missing = is.na(value), year = as.numeric(year)) %>% 
    group_by(countryName) %>% 
    summarise(n_nas  = sum(missing)) %>% 
    filter(n_nas > 0) %>% 
    mutate(countryName = as_factor(countryName) %>% fct_reorder(., n_nas)) %>% 
    ggplot(aes(countryName, n_nas)) + 
    geom_col(alpha = 0.5) + 
    geom_hline(yintercept = 10, color = "red", linetype = 2) +
    coord_flip() +
    theme_light(base_size = 7) + 
    labs(y = "Number of missing values", x = "Country", tag = "C") +
    theme(legend.position = c(0.8, 0.2))

g0 / (g1 + g2)

# ggsave(
#     filename = "ecofp_missing_values.png",
#     plot = (g0 / (g1 + g2)), 
#     device = "png", path = "figures/",
#     width = 7, height = 5, bg = "white", dpi = 400
# )
```

@@

```{r}
i2 <- inq %>% 
    filter(year > 1950, resource_detailed == "Income, gross") %>% 
    ggplot(aes(gini_reported, palma)) +
    geom_miss_point(size = 1, alpha = 0.5) +
    labs(y = "Palma index", x = "Gini index", tag = "B") + 
    scale_color_manual("Missing values", values = c("purple", "orange")) +
    theme_light(base_size = 6) +
    theme(legend.position = c(0.2,0.8))

i2
```

@@

```{r}
i1 <- inq %>% 
    filter(year > 1950, resource_detailed == "Income, gross") %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini") %>% 
    mutate(year = as.numeric(year)) %>% 
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = gini)) +
    scale_fill_viridis_c(
        "Mean Gini",option = "C", na.value = "grey75")+
    labs(x = "Year", y = "Countries (N=200)", tag = "A") +
    theme_light(base_size = 6) +
    theme(axis.text.y = element_blank(), legend.position = "top")

i1
```
```{r}
inq %>% 
    filter(!is.na(palma), year > 1990) %>% select(country, year, palma) %>% 
    group_by(country, year) %>% 
    summarize(palma_mean = mean(palma, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = palma_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "palma_mean") %>%
    mutate(year = as.numeric(year)) %>%
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = palma_mean)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```
To be fair with Gini, let's to the same over all Gini's reported. Even if Gini comes computed by different studies and in different currencies, the coefficient is always between 0-100, meaning it is comparable.

```{r}
i1 <- inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% 
    select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year)) %>%
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = gini_mean)) +
    scale_fill_viridis_c(
        "Mean Gini",option = "C", na.value = "grey75")+
    labs(x = "Year", y = "Countries (N=196)", tag = "A") +
    theme_light(base_size = 6) +
    theme(axis.text.y = element_blank(), legend.position = "top", 
          legend.key.height = unit(3, "mm"))
i1
```
So how many country / years can we use? Remember the data has `r inq$country %>% unique() %>% length()` countries

```{r}
i3 <- inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year), missing = is.na(gini_mean)) %>% 
    group_by(year) %>% 
    summarise(n_nas = sum(missing), missingness = n_nas / 200) |> 
    ggplot(aes(year, missingness)) +
    geom_col(alpha = 0.5) +
    theme_light(base_size = 6) +
    theme(legend.position = c(0.8,0.8)) +
    labs(y = "Proportion of countries with\n missing values", x="Year", tag = "C")

i3
```
We are dealing with ~50% of missing values on the best years, at best we can work with 100 countries. Which ones?

```{r}
i4 <- inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year), missing = is.na(gini_mean)) %>% 
    group_by(country) %>% 
    summarise(n_nas = sum(missing), missingness = n_nas / (2018-1990)) %>% 
    arrange(missingness) |> 
    filter(missingness <=0.5) |> 
    mutate(country = as_factor(country)) |> 
    ggplot(aes(missingness, country)) +
    geom_col(alpha = 0.5) +
    geom_vline(xintercept = 0.3, color = "red", linetype = 2) +
    theme_light(base_size = 6) + 
    labs(x = "Proportion of missing values", y = "Country", tag = "D") +
    theme(axis.text.y = element_text(size = 4))
i4
```

@@

```{r}
(i1 / (i2+i3+i4))

# ggsave(
#     filename = "inq_missing_values.png",
#     plot = (i1 / (i2+i3+i4)),
#     device = "png", path = "figures/",
#     width = 7, height = 6, bg = "white", dpi = 400
# )
```

@@

```{r}
dat1 %>% 
    select(-x35) %>% 
    filter(str_detect(indicator_name, "GNI")) %>% 
    pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>% 
    mutate(year = str_remove(year, "x")) %>% 
    mutate(year = as.numeric(year)) %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(color = country_name), show.legend = FALSE) +
    #geom_miss_point(show.legend = FALSE) +
    facet_wrap(~indicator_name, scales = "free_y")
```

```{r}
## How many missing values? 10506, but who has more?
missingness <- dat1 %>% 
    select(-x35) %>% 
    filter(str_detect(indicator_name, "GNI")) %>% 
    pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>% 
    mutate(year = str_remove(year, "x")) %>% 
    mutate(year = as.numeric(year)) %>% 
    mutate(ismissing = is.na(value)) %>% 
    group_by(country_code, indicator_name, country_name) %>% 
    summarize(missingness = sum(ismissing)/n())

p1 <- missingness %>% 
    ggplot(aes(country_code, indicator_name)) +
    geom_tile(aes(fill = missingness)) + 
    scale_fill_viridis_c(
        "Proportion of missing values", option = "D", na.value = "grey75")+
    labs(y = "Prosperity metrics", x = "Countries (N=239)", tag = "A") +
    theme_light(base_size = 6) +
    theme(axis.text.x = element_blank(), legend.position = "top", 
          legend.key.height = unit(3, "mm"))
p1

```

@@

```{r}
p2 <- missingness %>% 
    ungroup() %>% 
    group_by(indicator_name) %>% 
    summarize(x = mean(missingness)) %>% 
    arrange(desc(x)) |> 
    mutate(indicator_name = as_factor(indicator_name)) |> 
    ggplot(aes(x, indicator_name)) +
    geom_col(alpha = 0.5) +
    #geom_vline(xintercept = 0.3, color = "red", linetype = 2) +
    theme_light(base_size = 6) + 
    labs(x = "Proportion of missing values", y = "Prosperity metrics", tag = "B") +
    theme(axis.text.y = element_text(size = 6))
p2
```

```{r}
p3 <- missingness %>% 
    ungroup() %>% 
    group_by(country_name) %>% 
    summarize(x = mean(missingness)) %>% 
    arrange(x) |> 
    rownames_to_column("rank") |> 
    mutate(rank = as.numeric(rank)) |> 
    ggplot(aes(x, rank)) +
    geom_point(alpha = 0.5) +
    geom_vline(xintercept = 0.3, color = "red", linetype = 2) +
    theme_light(base_size = 6) + 
    labs(x = "Proportion of missing values", y = "Country rank", tag = "C") +
    theme(axis.text.y = element_text(size = 4))
p3
```
It seems that `GNI per capita (current US$)` is the one with less missing values. Tong recomments using the one with Atlas method wich is close on missingness, only 18%. A threshold of 50% missigness per country will left out places like Somalia, Iraq, Costa Rica, Zambia, China or Chile. Perhaps 60% is tolerable?

```{r}
p1 / (p2+p3)

ggsave(
    filename = "prosperity_missing_values.png",
    plot = p1 / (p2+p3),
    device = "png", path = "figures/",
    width = 7, height = 5, bg = "white", dpi = 400
)
```
